---
title             : "Developmental Changes of Sensitivty to Optic Flow from Childhood to Adulthood"
shorttitle        : "MOCO"
author: 
  - name          : "Rick O. Gilmore"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Psychology, The Pennsylvania State University, University Park, PA 16802 USA"
    email         : "rogilmore@psu.edu"
  - name          : "Yiming Qian"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Department of Psychology, The Pennsylvania State University"
  - id            : "2"
    institution   : "Databrary.org"
author_note: |
  The authors are with the Department of Psychology at The Pennsylvania State University.
  The authors acknowledge support from the Department of Psychology, the Social, Life, & Engineering Sciences Imaging Center (SLEIC), and the Child Study Center's Open Data in Developmental Science (ODDS) initiative.
abstract: |
  
  
keywords          : "Optic flow, visual perception, visual motion, coherence"
wordcount         : no
bibliography      : ["bib/references.bib"]
figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
params:
  data_file_in: '../data/survey_clean.csv'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, error=FALSE)
```

```{r load_packages, include = FALSE}
if(!require("papaja")){devtools::install_github("crsh/papaja")}
library("papaja")

if(!require("afex")){install.packages("afex")}
library("afex")

if(!require("gmodels")){install.packages("gmodels")}
library("gmodels")

# May need this for newer versions of R (needed for R 3.4.1)
# devtools::install_github("crsh/papaja")

if(!require("tidyverse")){install.packages("tidyverse")}
library("tidyverse")

# Could also source("R/Install_load_package.R") and then
# lapply(c("papaja", "tidyverse"), Install_load_package)
```

```{r analysis_preferences}
# Seed for random number generation
set.seed(42)
```
```{r import-survey}
#survey <- read_csv("data/survey_clean.csv")
# Or we can use "cleaned" actual survey data
survey <- read_csv(params$data_file_in)

survey$R_exp <- ordered(survey$R_exp, levels=c("none",
                                               "limited",
                                               "some",
                                               "lots",
                                               "pro"))

survey$Sub_id <- 1:length(survey$R_exp)
```

Author guidelines:
Developmetnal Psychology
Vision Research
Journal of Vision
Eye Perception
Perception of psychophysics
JEP: Human Perception
Visual Cognition
Perception



## Introduction

## Methods {#methods}
All participants had normal vision or corrected to normal vision. Written consent to participate was obtained from adult participants or from parents or guardians on behalf of child participants under  procedures approved by the Institutional Review Board of The Pennsylvania State University (#37946). The research was conducted according to the principles expressed in the Declaration of Helsinki. Participants were given credit or $10 per hour for their participation.

## Experiment 1: children
#### Participants
31 children (19  female; age range: 5.2-8.6 years; mean age: 6.7 years (SD=1.0)) participated in the study. The sample consisted of children drawn from a database of families in Centre County, Pennsylvania. Children were excluded if they were born prematurely, had a history of serious visual or medical problems, epilepsy, or seizures. One participant's data (1045) was excluded due to incomplete data collection.

#### Stimuli 

The stimuli was generated on an iMac computer using MATLAB (R2013b) and Psychtoolbox (XXX). The display consisted of two side-by-side, time varying (1.2 Hz coherent/incoherent cycle) annular-shaped (18 deg outer/5 deg inner diameter) optic flow displays with a fixation dot centrally located between the two circular random dot kinematogram displays. One display depicted random (0% coherent) motion while the other depicted a radial or linear motion pattern at one of four fixed coherence levels. Adult and child participants viewed different coherence level profiles. Four runs were presented to the participant in a single visit with 2 runs at a speed of 2 deg/s and 2 runs at a speed of 8 deg/s presented in a **balanced random order**. Each of the 8 combinations of coherence patterns and coherence levels appeared once on the left display and once on the right display for a total of 16 conditions within a block. A method of constant stimuli was used, and condition order varied randomly within a block. Five blocks were completed within a single run resulting in 80 trials per run. Each trial had a 10 second response time limit.

#### Procedures
Upon arrival at the laboratory, study and visit procedures were described and informed consent was obtained. Visual acuity was measured for each participant with the Snellen Visual Acuity chart. Participants were escorted into a dimly lit testing room and seated 60cm in front of the computer monitor. The experimenter remained in the testing room for the entire study.  Child observers fixated centrally and judged which side contained coherent motion, indicating their choice by verbalizing or pointing to the monitor. A practice session of 6-16 trials was used to teach the participants what to look for in the display for a given trial. The choice (left/right) was entered by the experimenter using a Sabrent game controller. Child participants viewed one of two coherence level profiles (20, 40, 60, 80%) or (15, 30, 45, 60%). The entire data collection session was video recorded. Participants were given the option to take a break after each run.

#### Data analysis
Each run generated an output file containing pattern type, coherence level, reaction time and accuracy for each trial. R (citation) scripts were were generated to clean and merge the data files prior to analysis. Statistical analysis and summary plots were generated using R Studio (citation). Repeated measures analysis of variance (ANOVA) was used to determine the effects of speed, pattern, and coherence. Prior to conducting statistical analyses, the individual percent correct by coherence, pattern and speed were plotted and visually evaluated. One adult participant (14) was eliminated from analysis for having chance or worse than chance performance even at the highest coherence levels.
### Results and discussion

## Experiment 2: Adults
### Method
#### Participants
30 adults (14 male, 16 female; age range: 18.7-23.9; mean age: 20.8 years, SD = X) participated in this study. The sample consisted of young adults recruited from the undergraduate psychology subject pool at the Pennsylvania State University.  One participant’s data was removed from analysis for failure to follow task instructions. Resulting in 29 adults 14 male, 15 female; age range: X; mean age: X years, SD = X) utilized for analysis.
#### Stimuli 
#### Procedures
Upon arrival at the laboratory, study and visit procedures were described and informed consent was obtained. Visual acuity was measured for each participant with the Snellen Visual Acuity chart. Participants were escorted into a dimly lit testing room and seated 60cm in front of the computer monitor. 
Participants were instructed to fixate on the dot in the center of the display and to use their peripheral vision to discern which of the displays was exhibiting an optic flow pattern. They were informed about the 10 second response limit and that their response times were being recorded. Participants used a keyboard to make their decisions, pressing the “z” key to choose the left and the “/ or ?” key to choose the right. The four coherence levels used were 5%, 10%, 15%, and 20%. Participants were given the option to take a break half way through the experiment.
#### Data analysis
Each run generated an output file containing pattern type, coherence level, reaction time and accuracy for each trial. R (citation) scripts were were generated to clean and merge the data files prior to analysis. Statistical analysis and summary plots were generated using R Studio (citation). Repeated measures analysis of variance (ANOVA) was used to determine the effects of speed, pattern, and coherence. Prior to conducting statistical analyses, the individual percent correct by coherence, pattern and speed were plotted and visually evaluated. One adult participant (14) was eliminated from analysis for having chance or worse than chance performance even at the highest coherence levels.
### Results and discussion

## General discussion

## Conclusion

## Reference
