---
title             : "Developmental Changes of Sensitivty to Optic Flow from Childhood to Adulthood"
shorttitle        : "MOCO"
author: 
  - name          : "Rick O. Gilmore"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Psychology, The Pennsylvania State University, University Park, PA 16802 USA"
    email         : "rogilmore@psu.edu"
  - name          : "Yiming Qian"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Department of Psychology, The Pennsylvania State University"
  - id            : "2"
    institution   : "Databrary.org"
author_note: |
  The authors are with the Department of Psychology at The Pennsylvania State University.
  The authors acknowledge support from the Department of Psychology, the Social, Life, & Engineering Sciences Imaging Center (SLEIC), and the Child Study Center's Open Data in Developmental Science (ODDS) initiative.
abstract: |
  
  
keywords          : "Optic flow, visual perception, visual motion, coherence"
wordcount         : no
bibliography      : ["bib/references.bib"]
figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
params:
  data_file_in: '../data/survey_clean.csv'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, error=FALSE)
```

```{r load_packages, include = FALSE}
if(!require("papaja")){devtools::install_github("crsh/papaja")}
library("papaja")

if(!require("afex")){install.packages("afex")}
library("afex")

if(!require("gmodels")){install.packages("gmodels")}
library("gmodels")

# May need this for newer versions of R (needed for R 3.4.1)
# devtools::install_github("crsh/papaja")

if(!require("tidyverse")){install.packages("tidyverse")}
library("tidyverse")

# Could also source("R/Install_load_package.R") and then
# lapply(c("papaja", "tidyverse"), Install_load_package)
```


Author guidelines:
Developmetnal Psychology
Vision Research
Journal of Vision
Eye Perception
Perception of psychophysics
JEP: Human Perception
Visual Cognition
Perception



## Introduction

## Methods {#methods}

### Participants
A total of 31 children between 5.2 and 8.6 years old (12 male, 19  female; mean age: 6.7 years ±1.0 years)) and 30 adults participated between 18.7 and 23.9 (14 male, 16 female; mean age: 20.8 years ± years ) participated in the study. The sample consisted of children drawn from a database of families in Centre County, Pennsylvania. Children were excluded if they were born prematurely, had a history of serious visual or medical problems, epilepsy, or seizures. One participant's data (1045) was excluded due to incomplete data collection. Adults participants consisted of young adults recruited from the undergraduate psychology subject pool at the Pennsylvania State University. One participant (14)’s data was removed from analysis for failure to follow task instructions. Resulting in 29 adults (14 male, 15 female; age range: X; mean age: X years, SD = X) utilized for analysis.

All participants had normal vision or corrected to normal vision. Written consent to participate was obtained from adult participants or from parents or guardians on behalf of child participants under  procedures approved by the Institutional Review Board of The Pennsylvania State University (#37946). The research was conducted according to the principles expressed in the Declaration of Helsinki. Participants were given credit or $10 per hour for their participation.

#### Stimuli 

The stimuli was generated on an iMac computer using MATLAB (R2013b) and Psychtoolbox (XXX). The display consisted of two side-by-side, time varying (1.2 Hz coherent/incoherent cycle) annular-shaped (18 deg outer/5 deg inner diameter) optic flow displays with a fixation dot centrally located between the two circular random dot kinematogram displays. One display depicted random (0% coherent) motion while the other depicted a radial or linear motion pattern at one of four fixed coherence levels. Adult and child participants viewed different coherence level profiles. Four runs were presented to the participant in a single visit with 2 runs at a speed of 2 deg/s and 2 runs at a speed of 8 deg/s presented in a **balanced random order**. Each of the 8 combinations of coherence patterns and coherence levels appeared once on the left display and once on the right display for a total of 16 conditions within a block. A method of constant stimuli was used, and condition order varied randomly within a block. Five blocks were completed within a single run resulting in 80 trials per run. Each trial had a 10 second response time limit.

#### Procedures
Upon arrival at the laboratory, study and visit procedures were described and informed consent was obtained. Visual acuity was measured for each participant with the Snellen Visual Acuity chart. Participants were escorted into a dimly lit testing room and seated 60cm in front of the computer monitor. 
With the child participants, the experimenter remained in the testing room for the entire study. Child observers fixated centrally and judged which side contained coherent motion, indicating their choice by verbalizing or pointing to the monitor. A practice session of 6-16 trials was used to teach the participants what to look for in the display for a given trial. The choice (left/right) was entered by the experimenter using a Sabrent game controller. Child participants viewed one of two coherence level profiles (20, 40, 60, 80%) or (15, 30, 45, 60%). The entire data collection session was video recorded. Participants were given the option to take a break after each run.

Adult Participants were instructed to fixate on the dot in the center of the display and to use their peripheral vision to discern which of the displays was exhibiting an optic flow pattern. They were informed about the 10 second response limit and that their response times were being recorded. Participants used a keyboard to make their decisions, pressing the “z” key to choose the left and the “/ or ?” key to choose the right. The four coherence levels used were 5%, 10%, 15%, and 20%. Participants were given the option to take a break half way through the experiment.

#### Data analysis
R (citation) scripts were were generated to clean and merge the data files prior to analysis. The trials with no reponse were coded as "FALSE" in accuracy and "NA"" in reaction time. Prior to conducting statistical analyses, the individual percent correct by coherence, pattern and speed were plotted and visually evaluated. One adult participant (14) was eliminated from analysis for having chance or worse than chance performance even at the highest coherence levels.
Statistical analysis and summary plots were generated using R Studio (citation). 

### Results 
A generalized linear mixed-effects model (GLMM) was used to investigate the effects of speed, pattern, and coherence on accuracy of direction discrimination task. Accuracy was predicted from the additive effects of age, coherence, speed, pattern type of optic flow and an interaction term between these terms, with a subject-level random intercept. A variety of links for the binomial family to fit m-alternative forced-choice psychophysical functions (probit, logit, cloglog, weibull) were used to find the best model fit. For both the child and adult participants, it was found GLMM with the logit link has the best fitting. In this model of child participants, it revealed statistical significant coefficients for gender, age, coherence, and optic flow pattern type, as well as the interaction between coherence and speed, speed and age, and pattern type and age. The interaction between age and coherence was excluded in the model although it was significant. This is because this interaction lied in the assumption of the experiment design (why?). As predicted, younger participants have lower accuracy than the older child participants. 5-year-old children has significantly lower accuracy than 7-year-old children (P < 0.001, effect size= ) and 8-year-old children (P < 0.001, effect size= ). Child participants have significant higher accuracy in the tasks of high coherence than in the low coherence (p<0.001, effect size=). Optic flow of radial pattern was found to have better detecton performance than linear pattern (P<0.001, effect size=). In addition, there is significant gender effect on accuracy. Male have better performance of the percentage of correct responses than female (P<0.05, effect size= ).Coh:Speed8deg/s p<0.05, Speed8deg/s:Age8,p<0.05; PatternTyperadial:Age6, p<0.01, PatternTyperadial:Age7, p<0.01. 
For adult participants, main effects of gender, coherence, optic flow speed and pattern type and interaction effect of coherence and pattern type were found in the percentage of correct responses. As predicted,  participants have significant higher accuracy in the tasks of high coherence than in the low coherence (p<0.001, effect size=). Optic flow of radial pattern was found to have poorer detecton performance than linear pattern (P<0.001, effect size=). Participants had higher accuracy for low-speed Optic flow than low-speed optic flow (P<0.001, effect size=). In addition, there is significant gender effect on accuracy. Male have better performance of the percentage of correct responses than female (P<0.05, effect size= ). Coh:PatternTyperadial p<0.001

A linear mixed-effects model (LMM) was used to investigate the possible predictors of reaction times. In this model, we found the residual of RT violated the assumption of normality in LMM. Thus, we transformed the dependent variable to be the log of reaction times. For the child participants,it revealed statistical significant coefficients for age, coherence, and optic flow pattern type and speed, as well as the interaction between speed and age. The interaction between age and coherence was excluded in the model although it was significant. There is no significant gender effect time on the reaction times. As predicted, younger participants have significantly longer reaction times than the older child participants (p<0.001, effect size=). Child participants have significant quicker responses in the tasks of high coherence than in the low coherence (p<0.001, effect size=). Optic flow of radial pattern was found to have better detecton performance than linear pattern (P<0.001, effect size=). Speed8deg/s:Age8,p<0.001;
In tha analysis of adult dataset,it was found statistical significant coefficients for coherence, and optic flow pattern type, as well as the interaction between speed and coherence, coherence and pattern type. There is no significant gender effect time on the reaction times. As predicted, participants have significant quicker responses in the tasks of high coherence than in the low coherence (p<0.001, effect size=). Optic flow of radial pattern was found to have better detecton performance than linear pattern (P<0.001, effect size=). Coh:Speed ,p<0.001;Coh:PatternType, p<0.01.

In the group comparison of adult participants and child participants, there are significant difference between adult and child participants in both accuracy (p<0.001, effect size= ) and reaction times performance (p<0.01, effect size= ). For accuracy, significant main effects of gender, optic flow coherence, speed and pattern type, as well as interaction effect between coherence and speed were found.  For reaction times, significant main effects of optic flow coherence, speed and pattern type, as well as interaction effect between optic flow coherence and pattern type were found. Besides, child participants are found have inferior performance (accuracy p< 0.001, effect size=; reaction time, p< 0.001, effect size) in low coherence level than adults but outperformed adults in high coherence level . Child participants were found have better performance facing low-speed optic flow but poorer performance in high-speed optic flow than adults (Accuracy, p< 0.05, effect size=; reaction time, p<0.001, effect size=).


## Discussion
There are several limitations in this study. First, although we controled to manipulate the same stimuli and experimental procedures to both adult and child participants, there were still some differences. The adult participants were experimented with lower-coherence stimuli due to their greater capacity of motion detection. Also, the child participants were were pointed to the direction and the experimenters recoded their responses, while the adult participants pressed the button in each trials. This may cause the delay of responses for child participants. 

## Conclusion

## Reference
