---
title: "Optic flow motion coherence task analyses and plots"
author: "Yiming Qian & Rick Gilmore"
date: "'r Sys.time()'"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

# Purpose

This document describes the analyses conducted on data from a behavioral study of perceptual sensitivity to optic flow patterns corrupted by random motion noise.

# Preliminary steps

Set chunk options, install libraries (if needed), load libraries, and set the random generator seed for reproducibility.
We set the chunk option `cache = TRUE` to save on computation time when we re-run the document.
Since the document is large and contains slow-to-execute model-fitting components, we suggest running the analysis and visualization chunks individually in sequence.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
# library(tidyverse)
# library(lme4)
# library(knitr)
# library(psyphy)
# library(sjPlot)
# library(dplyr)
# library(viridis)
# library(ggplot2)
# library(gridExtra)
# library(afex)
# library(emmeans)
# library(pwr)
# library(here)

if (!("pacman" %in% installed.packages()[,])) {
    install.packages("pacman")
  }
pacman::p_load(tidyverse, lme4, psyphy, sjPlot, viridis, gridExtra, afex, emmeans, pwr, here, neuropsychology)
# pacman::p_load(dplyr, ggplot2, knitr, psyphy, lme4, sjPlot)

set.seed(2020)
```

# Import aggregate data and clean

We presume that all commands are executed from the project root and that all file paths are relative to the project root: `root_dir`.

```{r}
root_dir <- here::here()
```

## Adult

```{r import-adult}
df_adult <- readr::read_csv("data/adult/aggregate/adult-aggregate.csv")
str(df_adult)
```

We summarize the mean percent correct in adults to check whether each participant has mean accuracy higher than 50% (guess rate).

```{r clean-adult}
df_adult$Speed <-
  factor(df_adult$Speed, labels = c("2 deg/s", "8 deg/s"))

df_adult$SubID <- factor(df_adult$SubID)

df_adult %>%
  group_by(SubID) %>%
  summarize(N.corr = sum(Acc),
            N.tot = n(),
            Pct.Corr = N.corr / N.tot) ->
  df_adult.summary

df_adult.summary
```

## Child

```{r import-child}
df_child <- readr::read_csv("data/child/aggregate/child-aggregate.csv")
str(df_child)
```


```{r clean-child}
# Convert speed to factor
df_child$Speed <-
  factor(df_child$Speed, labels = c("2 deg/s", "8 deg/s"))

# SubID to factor
df_child$SubID <- factor(df_child$SubID)

# Age to age in years as ordinal factor
df_child$Age <- df_child$AgeDays / 365.25
df_child$AgeYrs <- ordered(cut(
  df_child$AgeDays / 365.25,
  breaks = c(5, 6, 7, 8, 9),
  labels = c("5-year-old", "6-year-old", "7-year-old", "8-year-old")
))
```

```{r 25-child}
# Convert speed to factor
df_child<-df_child %>% filter(Group=="linear-radial coh [.15 .3 .45 .6]")

df_child %>%
  group_by(PatternType, Speed, Coh) %>%
  summarize(N.corr = sum(Acc),
            N.tot = n(),
            Pct.Corr = N.corr / N.tot) ->
  df_child.summary

df_child.summary

table(df_child$Gender)
table(df_child$Condition)
```

# Modeling

## Child GLMM

We perform GLMM on the child data sets to examine the effects of age, sex, coherence, speed and pattern type on accuracy responses. 
The dependent variable (DV) is the binary response, `Acc`, either right or wrong. 
The main independent variables (IVs) are age (`AgeDays`), sex (`gender`, 2 levels), coherence (`Coh`, 4 levels), speed (`Speed`, 2 levels), pattern type (`PatternType`, 2 levels). 
We seek the most parsimonious model with IVs that are interpretable. 
The following are the procedures that we use to find the best-fitting and most-parsimonious model:

(1) all IVs and their interactions
(2) three main experimental conditions (speed, coherence, pattern) and their interactions
(3) check the model with the significant effects
(4) add age and sex
(5) add sex-related and age-related interactions
(6) check the model with the significant effects
(7) compare three link function (weibull, logit, probit)

### All IVs and interactions

Check the GLMM with logit links functions with all IVs and their interactions.
*Change the chunk option `eval=TRUE` to evaluate this chunk when knitting the document or evaluate manually.*

```{r acc-logit-full-child, eval=FALSE}
form.fullmodel <-
  Acc ~ scale(Coh) * Speed * PatternType * scale(Age) * Gender + (1 |
                                                                    SubID)
model.acc.logit.full_c <-
  glmer(
    formula = form.fullmodel,
    family = binomial(mafc.logit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_child
  )
summary(model.acc.logit.full_c)
```

### Three experimental conditions + interactions

*Change the chunk option `eval=TRUE` to evaluate this chunk when knitting the document or evaluate manually.*

```{r acc-logit-reduced-child, eval=FALSE}
form.explor <- Acc ~ scale(Coh) * Speed * PatternType + (1 | SubID)
model.acc.logit.ex_c <-
  glmer(
    formula = form.explor,
    family = binomial(mafc.logit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_child
  )
summary(model.acc.logit.ex_c)
```

### Drop non-significant effects

*Change the chunk option `eval=TRUE` to evaluate this chunk when knitting the document or evaluate manually.*

```{r acc-logit-drop-interactions-child, eval=FALSE}
form.explor1 <-
  Acc ~ scale(Coh) + Speed + PatternType + scale(Coh) * Speed + PatternType*Speed + (1 |
                                                                   SubID)
model.acc.logit.ex1_c <-
  glmer(
    formula = form.explor1,
    family = binomial(mafc.logit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_child
  )
summary(model.acc.logit.ex1_c)
anova(model.acc.logit.ex1_c, model.acc.logit.ex_c)
```

### Add age and sex 

*Change the chunk option `eval=TRUE` to evaluate this chunk when knitting the document or evaluate manually.*

```{r acc-logit-drop-interactions-plus-age-plus-sex-child, eval=FALSE}
form.explor2 <-
  Acc ~ scale(Coh) + Speed + PatternType + Gender + scale(Age) + scale(Coh) *
  Speed + PatternType*Speed + (1 | SubID)
model.acc.logit.ex2_c <-
  glmer(
    formula = form.explor2,
    family = binomial(mafc.logit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_child
  )
summary(model.acc.logit.ex2_c)
anova(model.acc.logit.ex1_c, model.acc.logit.ex2_c)
```

### Add age-related and sex-related interactions 

*Change the chunk option `eval=TRUE` to evaluate this chunk when knitting the document or evaluate manually.*

```{r acc-logit-drop-interactions-plus-age-plus-sex-plus-interactions-child, eval=FALSE}
form_c.explor3 <-
  Acc ~ scale(Coh) + Speed + PatternType + scale(Age) + scale(Coh) *
  Speed + Speed * scale(Age) + PatternType * scale(Age) + Speed * Gender +
  PatternType * Gender + scale(Coh) *Gender+ (1 | SubID)
model.acc.logit.ex3_c <-
  glmer(
    formula = form_c.explor3,
    family = binomial(mafc.logit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_child
  )
summary(model.acc.logit.ex3_c)
car::Anova(model.acc.logit.ex3_c, model.acc.logit.ex2_c)
```

### Drop non-significant effects

*Change the chunk option `eval=TRUE` to evaluate this chunk when knitting the document or evaluate manually.*

```{r acc-logit-drop-interactions-plus-age-plus-sex-plus-interactions-reduced-child, eval=FALSE}
df_childx<-df_child
df_childx$Coh<-factor(df_childx$Coh )

form_c.explor4 <-
  Acc ~ Coh + Speed + PatternType + Gender + scale(Age) + Coh *
  Speed + Speed * scale(Age) +Coh *Gender+ (1 | SubID)
model.acc.logit_c <-
  glmer(
    formula = form_c.explor4,
    family = binomial(mafc.logit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_childx
  )
summary(model.acc.logit_c)

form_c.explor4 <-
  Acc ~ scale(Coh) + Speed + PatternType + Gender + scale(Age) + scale(Coh) *
  Speed + Speed * scale(Age) +scale(Coh) *Gender+ (1 | SubID)
model.acc.logit_c <-
  glmer(
    formula = form_c.explor4,
    family = binomial(mafc.logit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_child
  )
summary(model.acc.logit_c)
car::Anova(model.acc.logit_c, model.acc.logit.ex3_c)
```

### Check probit & Weibull link function

*Change the chunk option `eval=TRUE` to evaluate this chunk when knitting the document or evaluate manually.*

```{r check-probit-weibull-child, eval=FALSE}
model.acc.probit_c <-
  glmer(
    formula = form_c.explor3,
    family = binomial(mafc.probit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_child
  )
model.acc.weibull_c <-
  glmer(
    formula = form_c.explor3,
    family = binomial(mafc.cloglog(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_child
  )
car::Anova(model.acc.probit_c, model.acc.logit_c)
car::Anova(model.acc.weibull_c, model.acc.logit_c)
```

The logit link function has the best model fit.

### Further examine the final model

We calculate the chi-square, 95% CI, conditional R square, marginal R square.

```{r test-child}
car::Anova(model.acc.logit_c, type = 3)
# for binomial response data
#compute the marginal R-square
#compute the variance in the fitted values
VarF <-
  var(as.vector(fixef(model.acc.logit_c) %*% t(model.acc.logit_c@pp$X)))
# VarCorr() extracts variance components
# attr(VarCorr(lmer.model),’sc’)^2 extracts the residual variance, VarCorr()$plot extract the variance of the plot effect
VarF / (VarF + VarCorr(model.acc.logit_c)[[1]][1] + attr(VarCorr(model.acc.logit_c), 'sc') ^
          2 + (pi ^ 2) / 3)

#compute the conditionnal R-square
(VarF + VarCorr(model.acc.logit_c)[[1]][1]) / (VarF + VarCorr(model.acc.logit_c)[[1]][1] + (attr(VarCorr(model.acc.logit_c), 'sc') ^2) + (pi ^ 2) / 3)
# The residual deviance in logistic regression is fixed to (pi ^ 2) / 3

# get 95% CI
se <- sqrt(diag(vcov(model.acc.logit_c)))
# table of estimates with 95% CI
tab <-
  cbind(
    OR = fixef(model.acc.logit_c),
    LL = fixef(model.acc.logit_c) - 1.96 * se,
    UL = fixef(model.acc.logit_c) + 1.96 * se
  )
# however, this result is different from the below
print(exp(tab))
```
```{r post-hoc}
lsmeans::lsmeans(model.acc.logit_c, pairwise ~ Gender:Coh, adjust = "tukey")
```

## Adult GLMM

In this model, we perform GLMM on the adult datasets to examine the effects of sex, coherence, speed and pattern type on accuracy responses. 
The DV is the binary response of accuracy (`Acc`), either right or wrong. 
The main IVs are sex (`gender`, 2 levels), coherence (`Coh`, 4 levels), speed (`Speed`, 2 leveles), and pattern type (`PatternType`, 2 levels). 
We seek the most parsimous model with the IVs which can be interpreted. 
The procedures to find the model with the best model fits is very similar to those used to find the best-fitting child model:

(1) all IVs and their interactions
(2) three main experimental conditions (speed, coherence, pattern) and their interactions
(3) check the model with the significant effects
(4) add sex
(5) add sex-related interactions
(6) check the model with the significant effects
(7) compare three link function (Weibull, logit, probit)

### All IVs and interactions

```{r acc-logit-full-adult-1, eval=FALSE}
form.fullmodel <-
  Acc ~ scale(Coh) * Speed * PatternType * scale(Age) * Gender + (1 |
                                                                    SubID)
model.acc.logit.full_a <-
  glmer(
    formula = form.fullmodel,
    family = binomial(mafc.logit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_child
  )
summary(model.acc.logit.full_a)
```

### Three experimental conditions + interactions

```{r acc-logit-full-adult-2, eval=FALSE}
form.explor <- Acc ~ scale(Coh) * Speed * PatternType + (1 | SubID)
model.acc.logit.ex_a <-
  glmer(
    formula = form.explor,
    family = binomial(mafc.logit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_adult
  )
summary(model.acc.logit.ex_a)
```

### Drop non-significant effects

```{r acc-logit-full-adult-3, eval=FALSE}
form.explor1 <-
  Acc ~ scale(Coh) + Speed + PatternType + scale(Coh) * PatternType + (1 |
                                                                         SubID)
model.acc.logit.ex1_a <-
  glmer(
    formula = form.explor1,
    family = binomial(mafc.logit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_adult
  )
summary(model.acc.logit.ex1_a)
anova(model.acc.logit.ex1_a, model.acc.logit.ex_a)
```

### Add sex

```{r acc-logit-full-adult-4, eval=FALSE}
form.explor2 <-
  Acc ~ scale(Coh) + Speed + PatternType + scale(Coh) * PatternType + Gender + (1 |
                                                                                  SubID)
model.acc.logit.ex2_a <-
  glmer(
    formula = form.explor2,
    family = binomial(mafc.logit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_adult
  )
summary(model.acc.logit.ex2_a)
anova(model.acc.logit.ex1_a, model.acc.logit.ex2_a)
```

### Add sex-related interactions

Final model reported in manuscript.

```{r acc-logit-full-adult-5}
form_a.explor3 <-
  Acc ~ scale(Coh) + Speed + PatternType + scale(Coh) * PatternType + Gender + Gender *
  Speed + Gender * PatternType + (1 | SubID)
model.acc.logit_a <-
  glmer(
    formula = form_a.explor3,
    family = binomial(mafc.logit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_adult
  )
summary(model.acc.logit_a)
anova(model.acc.logit.ex2_a, model.acc.logit_a)
```

### Check probit & Weibull link function

```{r acc-logit-full-adult-6, eval=FALSE}
model.acc.probit_a <-
  glmer(
    formula = form_a.explor3,
    family = binomial(mafc.probit(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_adult
  )
model.acc.weibull_a <-
  glmer(
    formula = form_a.explor3,
    family = binomial(mafc.cloglog(2)),
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
    data = df_adult
  )
anova(model.acc.logit_a, model.acc.probit_a, model.acc.weibull_a)
```

The logit link function has the best model fit (lowest AIC, BIC, log likelihood, and deviance).

### Further examine the final model

we calculated the chi-square, 95% CI, conditional R square, marginal R square.

```{r test-adult}
car::Anova(model.acc.logit_a, type = 3)
# for binomial response data
#compute the marginal R-square
#compute the variance in the fitted values
VarF <-
  var(as.vector(fixef(model.acc.logit_a) %*% t(model.acc.logit_a@pp$X)))
# VarCorr() extracts variance components
# attr(VarCorr(lmer.model),’sc’)^2 extracts the residual variance, VarCorr()$plot extract the variance of the plot effect
VarF / (VarF + VarCorr(model.acc.logit_a)[[1]][1] + attr(VarCorr(model.acc.logit_a), 'sc') ^
          2 + (pi ^ 2) / 3)

#compute the conditionnal R-square
(VarF + VarCorr(model.acc.logit_a)[[1]][1]) / (VarF + VarCorr(model.acc.logit_a)[[1]][1] + (attr(VarCorr(model.acc.logit_a), 'sc') ^
                                                                                              2) + (pi ^ 2) / 3)
# The residual deviance in logistic regression is fixed to (pi ^ 2) / 3

# get 95% CI
se <- sqrt(diag(vcov(model.acc.logit_a)))
# table of estimates with 95% CI
tab <-
  cbind(
    OR = fixef(model.acc.logit_a),
    LL = fixef(model.acc.logit_a) - 1.96 * se,
    UL = fixef(model.acc.logit_a) + 1.96 * se
  )
# however, this result is different from the below
print(exp(tab))
```

### Post-hoc comparison interaction effects

```{r post-hoc}
lsmeans::lsmeans(model.acc.logit_a, pairwise ~ Gender:Speed, adjust = "tukey")
lsmeans::lsmeans(model.acc.logit_a, pairwise ~ Gender:PatternType, adjust =
                   "tukey")
```

# Plots 

## Theme for all figures

```{r ggplot-themes}
# "The APA has determined specifications for the size of figures and the fonts used in them. Figures of one column must be between 2 and 3.25 inches wide (5 to 8.45 cm). Two-column figures must be between 4.25 and 6.875 inches wide (10.6 to 17.5 cm). The height of figures should not exceed the top and bottom margins. The text in a figure should be in a san serif font (such as Helvetica, Arial, or Futura). The font size must be between eight and fourteen point. Use circles and squares to distinguish curves on a line graph (at the same font size as the other labels). "
theme.custom <- theme(panel.background = element_rect(fill = "white", colour = "grey50"),
                      plot.title = element_text(size=18, face="bold",hjust = 0.5),
                      axis.title.x = element_text(size=18),
                      axis.title.y = element_text(size=18),
                      strip.text = element_text(size=18),
                      axis.text = element_text(size=18),
                      legend.position = c(0.85,0.14),
                      legend.title=element_text(size=18),
#  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                   #   legend.title=element_blank(),
                      legend.text=element_text(size=18))
```

## Child figures

### Mean accuracy by pattern and age
```{r plot-acc-age-speed}
df_p3_c_acc_pat_age <- df_child %>%
  # filter(Group=="linear-radial coh [.15 .3 .45 .6]") %>%
  group_by(SubID, PatternType, Age) %>%
  summarize(
    N.corr = sum(Acc),
    N.tot = n(),
    Pct.Corr.mean = 100 * N.corr / N.tot
  )
p1 <- ggplot(data =  df_p3_c_acc_pat_age,
             aes(
               y = Pct.Corr.mean,
               x = Age,
               color = PatternType,
               shape = PatternType,
               group = PatternType
             )) +
  geom_point() +
  geom_smooth(method = "lm") +
  ggtitle("Children: Accuracy by age & pattern") +
  xlim(5, 9) +
  ylim(45, 100) +
  xlab("Age") +
  ylab("Mean Accuracy (%)") +
  scale_color_viridis(discrete = TRUE) +
  theme.custom +
  annotate(
    "text",
    x = 5,
    y = 100,
    label = "c)",
    size = 12
  )
p1
# ggsave("~/R/moco-psychophysics/analysis/pub_image/2_c_acc_speed_age_new.pdf", p2)
```


### Speed by coherence interaction

```{r}
# df_child$Coh<-factor(df_child$Coh)
df_p2_c_acc_speed_coh <- df_child %>%
  group_by(SubID, Speed, Coh) %>%
  summarize(N.corr = sum(Acc),
            N.tot = n(),
            Pct.Corr = N.corr / N.tot)  %>%
  group_by(Speed, Coh) %>%
  mutate(
    n = length(Pct.Corr),
    Pct.Corr.mean = 100 * mean(Pct.Corr, na.rm = TRUE),
    Pct.Corr.ci = 1.96 * 100 * sd(Pct.Corr, na.rm = TRUE) / sqrt(length(Pct.Corr))
  )  #computation of the standard error of the mean, ci is 1.96*sem
df_p2_c_acc_speed_coh$Coh <-
  factor(
    df_p2_c_acc_speed_coh$Coh,
    levels = c("0.15", "0.3", "0.45", "0.6"),
    labels = c("15%", "30%", "45%", "60%")
  )
p3 <-
  ggplot(df_p2_c_acc_speed_coh,
         aes(
           x = Coh,
           y = Pct.Corr * 100,
           color = Speed,
           shape = Speed
         ),
         alpha = 0.5) +
  geom_point(
    data = df_p2_c_acc_speed_coh,
    aes(x = Coh, y = Pct.Corr.mean, color = Speed),
    size = 3,
    position = position_dodge(0.8)
  ) +
  geom_errorbar(
    aes(
      ymin = Pct.Corr.mean - Pct.Corr.ci,
      ymax = Pct.Corr.mean + Pct.Corr.ci,
      color = Speed
    ),
    position = position_dodge(0.8),
    width = .5
  )  +
  geom_jitter(position = position_dodge(0.8)) +
  ggtitle("Children: Accuracy by coherence & speed") +
  xlab("Coherence") +
  ylim(45, 100) +
  theme.custom + # Position legend in bottom right
  ylab("Mean Accuracy (%)") +
  scale_color_viridis(discrete = TRUE) +
  annotate(
    "text",
    x = 0.6,
    y = 100,
    label = "b)",
    size = 12
  ) 
#  theme(legend.title=element_blank())
p3
# ggsave("~/R/moco-psychophysics/analysis/pub_image/8_c_acc_speed_coh_new.pdf", p1)
```

### Speed by age interaction

```{r plot-acc-age-speed}
df_p3_c_acc_speed_age <- df_child %>%
  # filter(Group=="linear-radial coh [.15 .3 .45 .6]") %>%
  group_by(SubID, Speed, Age) %>%
  summarize(
    N.corr = sum(Acc),
    N.tot = n(),
    Pct.Corr.mean = 100 * N.corr / N.tot
  )
p2 <- ggplot(data =  df_p3_c_acc_speed_age,
             aes(
               y = Pct.Corr.mean,
               x = Age,
               color = Speed,
               shape = Speed,
               group = Speed
             )) +
  geom_point() +
  geom_smooth(method = "lm") +
  ggtitle("Children: Accuracy by age & speed") +
  xlim(5, 9) +
  ylim(45, 100) +
  xlab("Age") +
  ylab("Mean Accuracy (%)") +
  scale_color_viridis(discrete = TRUE) +
  theme.custom +
  annotate(
    "text",
    x = 5,
    y = 100,
    label = "c)",
    size = 12
  )
p2
# ggsave("~/R/moco-psychophysics/analysis/pub_image/2_c_acc_speed_age_new.pdf", p2)
```

### Mean accuracy by gender and coh

```{r}
df_p2_a_acc_gen_coh <- df_child %>%
  group_by(SubID, Coh, Gender) %>%
  summarize(N.corr = sum(Acc),
            N.tot = n(),
            Pct.Corr = N.corr / N.tot)  %>%
  group_by(Coh, Gender) %>%
  mutate(
    n = length(Pct.Corr),
    Pct.Corr.mean = 100 * mean(Pct.Corr, na.rm = TRUE),
    Pct.Corr.ci = 1.96 * 100 * sd(Pct.Corr, na.rm = TRUE) / sqrt(length(Pct.Corr))
  )  #computation of the standard error of the mean, ci is 1.96*sem
df_p2_a_acc_gen_coh$Coh <-
  factor(
    df_p2_a_acc_gen_coh$Coh,
    levels = c("0.15", "0.3", "0.45", "0.6"),
    labels = c("15%", "30%", "45%", "60%")
  )
p4 <-
  ggplot(
    df_p2_a_acc_gen_coh,
    aes(
      x = Coh,
      y = Pct.Corr * 100,
      color = Gender,
      shape = Gender
    ),
    alpha = 0.5
  ) +
  geom_point(
    data = df_p2_a_acc_gen_coh,
    aes(x = Coh, y = Pct.Corr.mean, color = Gender),
    size = 3,
    position = position_dodge(0.8)
  ) +
  geom_errorbar(
    aes(
      ymin = Pct.Corr.mean - Pct.Corr.ci,
      ymax = Pct.Corr.mean + Pct.Corr.ci,
      color = Gender
    ),
    position = position_dodge(0.8),
    width = .5
  )  +
  geom_jitter(position = position_dodge(0.8)) +
  ggtitle("Children: Accuracy by sex & coherence") +
  theme.custom + # Position legend in bottom right
  ylab("Mean Accuracy (%)") +
  ylim(45, 100) +
  scale_color_viridis(discrete = TRUE) +
  annotate(
    "text",
    x = 0.5,
    y = 100,
    label = "d)",
    size = 12
  )
p4
```

### Combine figures

```{r}
#p = grid.arrange(ncol = 2, nrow = 2, p0, p1, p2,p3)
p_c = grid.arrange(layout_matrix = rbind(c(1, 2),
                                         c(3, 4)),  p1, p2, p3, p4)
#labels = c("a", "b", "c")
ggsave(
  plot = p_c,
  filename = paste0(here::here(), "/analysis/pub_image/child1.png"),
  units = "in",
  height = 12,
  width = 12,
  dpi = 600
)
```

## Adult figures

### Mean accuracy by speed and age 

```{r plot-acc-age-speed}
df_p4_c_acc_speed_age <- df_adult %>%
  # filter(Group=="linear-radial coh [.15 .3 .45 .6]") %>%
  group_by(SubID, Speed, Age) %>%
  summarize(
    N.corr = sum(Acc),
    N.tot = n(),
    Pct.Corr.mean = 100 * N.corr / N.tot
  )
p3 <- ggplot(data = df_p4_c_acc_speed_age,
             aes(
               y = Pct.Corr.mean,
               x = Age,
               color = Speed,
               shape = Speed,
               group = Speed
             )) +
  geom_point() +
  geom_smooth(method = "lm") +
  ggtitle("Adults: Percent correct by coherence & pattern") +
  #  xlim(18,24) +
  xlab("Age") +
  ylab("Mean Accuracy (%)") +
  scale_color_viridis(discrete = TRUE) +
  theme.custom
p3
# ggsave("~/R/moco-psychophysics/analysis/pub_image/3_a_acc_speed_age_new.pdf", p3)
```

### Pattern by coherence interaction

```{r}
df_p2_a_acc_pat_coh <- df_adult %>%
  group_by(SubID, PatternType, Coh) %>%
  summarize(N.corr = sum(Acc),
            N.tot = n(),
            Pct.Corr = N.corr / N.tot)  %>%
  group_by(PatternType, Coh) %>%
  mutate(
    n = length(Pct.Corr),
    Pct.Corr.mean = 100 * mean(Pct.Corr, na.rm = TRUE),
    Pct.Corr.ci = 1.96 * 100 * sd(Pct.Corr, na.rm = TRUE) / sqrt(length(Pct.Corr))
  )  #computation of the standard error of the mean, ci is 1.96*sem
df_p2_a_acc_pat_coh$Coh <-
  factor(
    df_p2_a_acc_pat_coh$Coh,
    levels = c("0.05", "0.1", "0.15", "0.2"),
    labels = c("5%", "10%", "15%", "20%")
  )
df_p2_a_acc_pat_coh$Pattern <-
  factor(
    df_p2_a_acc_pat_coh$PatternType,
    levels = c("linear", "radial"),
    labels = c("Linear", "Radial")
  )
p4 <-
  ggplot(
    df_p2_a_acc_pat_coh,
    aes(
      x = Coh,
      y = Pct.Corr * 100,
      color = Pattern,
      shape = Pattern
    ),
    alpha = 0.1
  ) +
  geom_point(
    data = df_p2_a_acc_pat_coh,
    aes(x = Coh, y = Pct.Corr.mean, color = Pattern),
    size = 3,
    position = position_dodge(0.8)
  ) +
  geom_errorbar(
    aes(
      ymin = Pct.Corr.mean - Pct.Corr.ci,
      ymax = Pct.Corr.mean + Pct.Corr.ci,
      color = Pattern
    ),
    position = position_dodge(0.8),
    width = .5
  )  +
  geom_jitter(position = position_dodge(0.8)) +
  ggtitle("Adults: Percent correct by coherence & pattern") +
  xlab("Coherence") +
  theme.custom + # Position legend in bottom right
  ylim(30, 100) +
  ylab("Mean Accuracy (%)") +
  scale_color_viridis(discrete = TRUE) +
  annotate(
    "text",
    x = 0.6,
    y = 100,
    label = "a)",
    size = 12
  )
p4
# ggsave("~/R/moco-psychophysics/analysis/pub_image/8_a_acc_pattern_coh_new.pdf", p4)
```

### Speed by sex interaction

```{r}
df_p2_a_acc_gen_speed <- df_adult %>%
  group_by(SubID, Speed, Gender) %>%
  summarize(N.corr = sum(Acc),
            N.tot = n(),
            Pct.Corr = N.corr / N.tot)  %>%
  group_by(Speed, Gender) %>%
  mutate(
    n = length(Pct.Corr),
    Pct.Corr.mean = 100 * mean(Pct.Corr, na.rm = TRUE),
    Pct.Corr.ci = 1.96 * 100 * sd(Pct.Corr, na.rm = TRUE) / sqrt(length(Pct.Corr))
  )  #computation of the standard error of the mean, ci is 1.96*sem
# df_p2_a_acc_gen_speed$Speed<-factor(df_p2_a_acc_gen_speed$Speed, levels=c("2 deg/s","8 deg/s"), labels=c("2 deg/s","8 deg/s"))
p5 <-
  ggplot(
    df_p2_a_acc_gen_speed,
    aes(
      x = Speed,
      y = Pct.Corr * 100,
      color = Gender,
      shape = Gender
    ),
    alpha = 0.5
  ) +
  geom_point(
    data = df_p2_a_acc_gen_speed,
    aes(x = Speed, y = Pct.Corr.mean, color = Gender),
    size = 3,
    position = position_dodge(0.8)
  ) +
  geom_errorbar(
    aes(
      ymin = Pct.Corr.mean - Pct.Corr.ci,
      ymax = Pct.Corr.mean + Pct.Corr.ci,
      color = Gender
    ),
    position = position_dodge(0.8),
    width = .5
  )  +
  geom_jitter(position = position_dodge(0.8)) +
  ggtitle("Adults: Percent correct by sex & speed") +
  theme.custom + # Position legend in bottom right
  ylab("Mean Accuracy (%)") +
  ylim(30, 100) +
  scale_color_viridis(discrete = TRUE)  +
  annotate(
    "text",
    x = 0.5,
    y = 100,
    label = "b)",
    size = 12
  )
p5
# ggsave("~/R/moco-psychophysics/analysis/pub_image/9_a_acc_speed_gender_new.pdf", p5)
```

### Pattern by sex interaction

```{r}
df_p2_a_acc_gen_pat <- df_adult %>%
  group_by(SubID, PatternType, Gender) %>%
  summarize(N.corr = sum(Acc),
            N.tot = n(),
            Pct.Corr = N.corr / N.tot)  %>%
  group_by(PatternType, Gender) %>%
  mutate(
    n = length(Pct.Corr),
    Pct.Corr.mean = 100 * mean(Pct.Corr, na.rm = TRUE),
    Pct.Corr.ci = 1.96 * 100 * sd(Pct.Corr, na.rm = TRUE) / sqrt(length(Pct.Corr))
  )  #computation of the standard error of the mean, ci is 1.96*sem
df_p2_a_acc_gen_pat$Pattern <-
  factor(
    df_p2_a_acc_gen_pat$PatternType,
    levels = c("linear", "radial"),
    labels = c("Linear", "Radial")
  )
p6 <-
  ggplot(
    df_p2_a_acc_gen_pat,
    aes(
      x = Pattern,
      y = Pct.Corr * 100,
      color = Gender,
      shape = Gender
    ),
    alpha = 0.5
  ) +
  geom_point(
    data = df_p2_a_acc_gen_pat,
    aes(x = Pattern, y = Pct.Corr.mean, color = Gender),
    size = 3,
    position = position_dodge(0.8)
  ) +
  geom_errorbar(
    aes(
      ymin = Pct.Corr.mean - Pct.Corr.ci,
      ymax = Pct.Corr.mean + Pct.Corr.ci,
      color = Gender
    ),
    position = position_dodge(0.8),
    width = .5
  )  +
  geom_jitter(position = position_dodge(0.8)) +
  ggtitle("Adults: Percent correct by sex & pattern") +
  theme.custom + # Position legend in bottom right
  ylab("Mean Accuracy (%)") +
  ylim(30, 100) +
  scale_color_viridis(discrete = TRUE) +
  annotate(
    "text",
    x = 0.5,
    y = 100,
    label = "c)",
    size = 12
  )
p6
# ggsave("~/R/moco-psychophysics/analysis/pub_image/8_a_acc_pattern_gender_new.pdf", p6, height=7,width=7)
```

### Combine plots into one 

```{r}
# combine multiple plots into one
# pdf("~/R/moco-psychophysics/analysis/pub_image/adult.pdf")

#set the create the frames
# layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
# or
# par(mfcol=c(2,2))
#dev.off shuts down the specified (by default the current) graphical device
#here it passes the picture to the file
# dev.off()

#p = grid.arrange(ncol = 2, nrow = 2, p4, p5, p6)
p = grid.arrange(layout_matrix = rbind(c(NA, 1, 1, NA),
                                       c(2, 2, 3, 3)), p4, p5, p6)
#labels = c("a", "b", "c")
ggsave(
  plot = p,
  filename = paste0(here::here(), "/analysis/pub_image/adult1.png"),
  units = "in",
  height = 12,
  width = 12,
  dpi = 600
)
```

# Signal detection theory (SDT) analysis

We conducted a comparison between children and adults at the coherence level of 15% using Signal Detection Theory.

### Prepare data

First, combine these two datasets and calculate the d' and c in the child and adult datasets from SDT.

```{r get-SDT}
# library(psycho)
# Let's simulate three participants with different results at a perceptual detection task
df_child2 <- df_child %>%
  dplyr::select(SubID, PatternType, RT, Acc, Speed, Gender, Coh, Response, Age) %>%
  dplyr::filter(Coh == 0.15) %>%
  dplyr::filter(Response %in% c("L", "R"))
str(df_child2)

df_adult2 <- df_adult %>%
  dplyr::select(SubID, PatternType, RT, Acc, Speed, Gender, Coh, Response, Age) %>%
  dplyr::filter(Coh == 0.15)  %>%
  dplyr::filter(Response %in% c("L", "R"))
df_adult2$SubID <- factor(df_adult2$SubID)
str(df_adult2)

df2 <- rbind(df_child2, df_adult2)
df.new <- df2 %>%
  mutate(group = c(rep('child', nrow(df_child2)), rep('adult', nrow(df_adult2)))) %>%
  filter(Response != "NA") %>%
  group_by(SubID, group, Coh, PatternType, Speed, Gender) %>%
  summarize(
    n_hit = sum(Response == "L" & Acc == "TRUE"),
    n_fa  = sum(Response == "L" & Acc == "FALSE"),
    n_miss = sum(Response == "R" & Acc == "FALSE"),
    n_cr = sum(Response == "R" & Acc == "TRUE")
  )

indices <-
  neuropsychology::dprime(df.new$n_hit, df.new$n_fa, df.new$n_miss, df.new$n_cr)
df.sdt <- cbind(df.new, indices)
```

### ANOVA

We conduct four-way ANOVA to examine the effect of sex, group, speed and pattern on perceptual sensivitity (d') and include post hoc comparisons. 

``` {r anova-SDT-d'}
df.sdt$SubID <- as.integer(df.sdt$SubID)
a_d <-
  aov_ez(
    id = "SubID",
    dv = "dprime",
    df.sdt,
    between = c("Gender", "group"),
    within = c("Speed", "PatternType")
  )
summary(a_d)
knitr::kable(nice(a_d))
# calculate power
library(pwr)
pwr.anova.test(
  k = 2,
  n = 50,
  f = 0.17,
  sig.level = 0.05,
  power = NULL
)  # for pattern
pwr.anova.test(
  k = 2,
  n = 29,
  f = 0.15,
  sig.level = 0.05,
  power = NULL
)  # for group
pwr.f2.test(
  u = 27,
  v = 1,
  f2 = 34.94,
  sig.level = 0.05,
  power = NULL
) # age effect in children
## chilren only
df.sdt.c <- df.sdt %>% filter(group == "child")
a_d.c <-
  aov_ez(
    id = "SubID",
    dv = "dprime",
    df.sdt.c,
    between = c("Gender"),
    within = c("Speed", "PatternType")
  )
summary(a_d.c)
knitr::kable(nice(a_d.c))
em_c <- emmeans(a_d.c, ~ Gender * PatternType)
pairs(em_c)
em_c1 <- emmeans(a_d.c, ~ Gender * Speed)
pairs(em_c1)
## adults only
df.sdt.a <- df.sdt %>% filter(group == "adult")
a_d.a <-
  aov_ez(
    id = "SubID",
    dv = "dprime",
    df.sdt.a,
    between = c("Gender"),
    within = c("Speed", "PatternType")
  )
summary(a_d.a)
knitr::kable(nice(a_d.a))
em_a <- emmeans(a_d.a, ~ Gender * PatternType)
pairs(em_a)
em_a1 <- emmeans(a_d.a, ~ Gender * Speed)
pairs(em_a1)

# What about the estimated marginal means of the design?
emmeans(a_d, ~ Gender * group * Speed * PatternType)
# What about the ‘main effects?
em1 <- emmeans(a_d, ~ Gender)
em2 <- emmeans(a_d, ~ group)
em3 <- emmeans(a_d, ~ Speed)
em4 <- emmeans(a_d, ~ PatternType)
em5 <- emmeans(a_d, ~ Speed * group)
em6 <- emmeans(a_d, ~ PatternType * group)
em7 <- emmeans(a_d, ~ Gender * PatternType)
em8 <- emmeans(a_d, ~ Gender * PatternType * group)
# And pairwise difference among levels of inference validity?
em1
pairs(em1)
df.sdt %>%
  dplyr::group_by(group) %>%
  summarize(mean = mean(dprime))
```

### Plt d' in two conditions of pattern and speed.

``` {r anova-plot}
# Plotting results of aov_ez
emmip(a_d, group ~ PatternType|Speed, CIs = TRUE)
```