---
title: "moco-adult-laminar-radial-summary"
output: html_document
---

```{r}
dts=format( Sys.time(), "%Y-%m-%d-%H%M%S")
sprintf("Analysis conducted %s.", dts)
```

## Source functions & load packages
```{r}
source("bulk.import.convert.sessions.R")
library(ggplot2)
library(dplyr)
```

## Data importing and clean-up

### Import and convert all data
```{r}
b <- bulk.import.convert.sessions()
moco <- read.csv("../aggregate-data/adult-laminar-radial-grouped.csv")
```

### Convert some variables to more workable forms
```{r}
moco$SessionDate = as.Date( moco$SessionDate )
moco$DegPSec = ordered( moco$DegPSec )
moco$Acc = as.logical( moco$Acc )
moco$ParticipantID = as.factor( moco$ParticipantID )
```

### Summarize imported data
```{r}
summary( moco )
```

### Convert to dplyr tableand generate some summaries
```{r}
tbl.moco <- tbl_df( moco )
```

### Plot P.Corr by Pattern and Speed
```{r}
tbl.p.corr <- tbl.moco %>%
  group_by( DegPSec, PatternType, Coh, ParticipantID ) %>%
  summarise( P.Corr=mean( Acc, na.rm=TRUE ), ntrials=n() )
tbl.p.corr

# Plot R by DegPSec and PatternType
pl.p.corr <- ggplot( data=tbl.p.corr ) +
  facet_grid( facets= DegPSec ~ PatternType ) +
  geom_point( aes(x=Coh, y=P.Corr, color=ParticipantID ) ) +
  geom_line( aes(x=Coh, y=P.Corr, group=ParticipantID, color=ParticipantID) )

pl.p.corr

dts=format( Sys.time(), "%Y-%m-%d-%H%M%S")
fn=sprintf("%s/%s-%s.png", "../imgs", dts, "p.corr")
ggsave( file=fn ) 
```

### Plot RT by Pattern & Speed
```{r}
tbl.rt <- tbl.moco %>%
  group_by( DegPSec, PatternType, Coh, ParticipantID ) %>%
  summarise( RT.mean=mean( RT, na.rm=TRUE ), 
             RT.sem = sd(RT)/sqrt( n() ) )

tbl.rt

limits = aes( ymax = RT.mean + RT.sem, ymin = RT.mean - RT.sem, group=ParticipantID )

pl.rt <- ggplot( data=tbl.rt, aes(x=Coh, y=RT.mean, group=ParticipantID, color=ParticipantID) ) +
  facet_grid( facets = DegPSec ~ PatternType ) +
  geom_line() +
  geom_pointrange( limits )

pl.rt

dts=format( Sys.time(), "%Y-%m-%d-%H%M%S")
fn=sprintf("%s/%s-%s.png", "../imgs", dts, "rt")
ggsave( file=fn ) 

```

### Speed/Accuracy Trade-off
```{r}
tbl.by.sub <- tbl.moco %>%
  group_by( DegPSec, PatternType, Coh, ParticipantID ) %>%
  summarise( P.Corr = mean( Acc, na.rm=TRUE ), RT.mean=mean(RT, na.rm=TRUE), RT.sem=sd(RT, na.rm=TRUE)/sqrt(n()) )

pl.speed.accuracy <- ggplot( tbl.by.sub ) +
  geom_line( aes(x=P.Corr, y=RT.mean, shape=as.factor(Coh), color=ParticipantID, group=ParticipantID)) +
  geom_pointrange( aes( x=P.Corr, y=RT.mean, ymax = RT.mean + RT.sem, ymin=RT.mean-RT.sem, group=ParticipantID, shape=as.factor(Coh), color=ParticipantID ) ) + 
  facet_grid( facets = DegPSec ~ PatternType )

pl.speed.accuracy

dts=format( Sys.time(), "%Y-%m-%d-%H%M%S")
fn=sprintf("%s/%s-%s.png", "../imgs", dts, "spd.acc")
ggsave( file=fn ) 
```

Note possible steeper slope for slow (2 deg/s) laminar flows. Keep on the look out for this and analyze formally later.